---
title: "regression_analysis_prageeth"
author: "my author"
date: "2025-02-25"
output: beamer_presentation
---
#
#```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = FALSE)
#```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```








# Start my work.

##
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

chooseCRANmirror(graphics = FALSE, ind = 1)


packages <- c("ggplot2", "mgcv", "MASS", "caret", "knitr")
install_if_missing <- packages[!(packages %in% installed.packages()[,"Package"])]
if(length(install_if_missing)) install.packages(install_if_missing)


library(ggplot2)
library(mgcv)    # Generalized Additive Models (GAM)
library(MASS)    # Stepwise regression
library(caret)   # Train-test split, cross-validation
library(knitr)   # For mark
```



```{r}
house_df <- read.csv("C:\\Users\\Prageeth\\Source\\MSC\\seminar-course\\advanced-regression-kc\\kc_house_data.csv", stringsAsFactors = TRUE)
features <- c("bedrooms", "bathrooms", "sqft_living", "sqft_lot", "floors", 
              "condition", "grade", "sqft_above", "sqft_basement", "yr_built")
target <- "price"
```


# Simple Linear Regression
##
```{r,echo = TRUE,eval=FALSE,fig.width=3, fig.height=5}
simple_lm <- lm ( price ~ sqft_living ,
                  data = house_df )
ggplot(house_df, aes(x = sqft_living, y = price)) +
  geom_point(alpha = 0.5) +  # Scatter plot
  geom_smooth(method = "lm", color = "blue", se = FALSE) +# Regression line
  labs(title = "Price vs. Size Fit",
       x = "Size (sqft_living)",
       y = "Price") +
  theme_minimal()
  
simple_lm
```


#
## output

```{r,echo=FALSE,fig.width=5, fig.height=1.5}
simple_lm <- lm ( price ~ sqft_living ,
                  data = house_df )
                  
                  
ggplot(house_df, aes(x = sqft_living, y = price)) +
  geom_point(alpha = 0.5) +  # Scatter plot
  geom_smooth(method = "lm", color = "blue", se = FALSE) +  # Regression line
  labs(title = "Price vs. Size Fit",
       x = "Size (sqft_living)",
       y = "Price") +
  theme_minimal()
  
  
```

```{r,echo=FALSE}

simple_lm
```


#
##Model Summary (Extracts Coefficients, p-values, R²)

```{r,echo=FALSE, width=3.5}
house_lm <- lm(price ~ sqft_living + sqft_lot + bathrooms + grade, 
               data = house_df, na.action = na.omit)
summary(house_lm)
```

# 


# Extract RMSE
```{r,echo=FALSE}
predictions <- predict(house_lm, newdata = house_df)
residuals <- house_df$price - predictions
RMSE <- sqrt(mean(residuals^2))
```

# Extract R-squared
```{r,echo=FALSE}
R_squared <- summary(house_lm)$r.squared
```

# Extract P-values
```{r,echo=FALSE}
p_values <- summary(house_lm)$coefficients[, 4]
```

# Print Results
```{r,echo=FALSE}
cat("RMSE:", RMSE, "\n")
cat("R²:", R_squared, "\n")
cat("P-values:\n")
print(p_values)
```

# Print Results
```{r,echo=FALSE}
house_lm_2 <- lm(price ~ sqft_living + sqft_lot15 + bathrooms +
                 bedrooms + grade, data=house_df, na.action=na.omit)
house_lm_2


house_poly <- lm(price ~ poly(sqft_living, 2) + sqft_lot15 + bathrooms +
                   bedrooms + grade, data=house_df)
house_poly

```